### 2014-11-11 tex/related.tex
Started related work section. Added macro for GSE/GSESH to paper.tex. Updated bibliography

The related work goes here.

>>>>

\subsection{Symbolic Exection}
Generalized symbolic execution \cite{GSE03} was the first symbolic execution method to handle linked data structures as inputs. Subsequent improvements to the method \cite{Deng:2006,Deng:2007} reduce path explosion by delaying path splitting over aliasing choices for as long as possible. However, since all of the above methods introduce nondeterminism during field reads and writes, they all worsen the path explosion problem to some extent or another. Since our heap representation does not introduce any nondeterminism beyond the level in standard symbolic execution, it will generally create fewer paths than these methods, sometimes by a wide margin.

\subsection{Model Checking}
Need to talk about Bounded Model Checking, etc.

\subsection{Static Analysis}
A number of static analysis techniques have been proposed that have some ability to reason about input heap structures. These techniques are usually over-approximations, often have difficulty reasoning about path-sensitive properties, or cannot reason about programs with recursive data structures. Moreover, static analysis methods are seldom useful for test-case generation.

One notable static analysis method that is closely related to ours is the procedure summaries proposed by Isil and Tomas Dillig \cite{Dillig:2011}. Elements of their technique that are similar to ours include; references guarded by constraint expressions, conditional write semantics, and a general aversion to path splitting. However, their method is limited by most of the standard problems with static analysis techniques, including path sensitivity, imprecision, and recursive data structures. While their method has improved path sensitivity over standard static analysis methods, it remains limited by a statically-generated symbolic input heap, and cannot adapt to path-sensitive heap properties of the input heap. Thus, their method cannot determine a priori which nodes in the input heap are actually accessed by the program. It may potentially generate input heaps containing a significant number of nodes that have no impact on program execution.

\subsection{Heap Representations}
A number of other techniques model references using sets of guarded locations \cite{Xie:2005,Cherem:2007,Dillig:2011}. These sets, sometimes known as value sets or value summaries, allow multiple heaps to be represented simultaneously with a higher degree of precision than afforded by shape analysis. However, some of these previous attempts utilize non-local recursive definitions for objects. The impact of using recursive definitions for objects is two-fold. First, these techniques must use approximations to resolve reference inequalities, resulting in false positives in the results. Second, these techniques are incapable of reasoning about heaps with loops in the data structures due to infinite recursion during dereferencing. This means that either a large class of data structures must be avoided, or a large class of behaviors must be ignored. In our heap representation, object descriptions are entirely local, thus avoiding these pitfalls.

### 2014-11-11 tex/results.tex
Started related work section. Added macro for GSE/GSESH to paper.tex. Updated bibliography

\section{Empirical Evaluation}
GSESH creates fewer execution paths than GSE, but it does so with increased path constraint complexity. Determining the impact of this tradeoff in real-world applications is an important research question.

Hypothesis

\subsection{Implementation}

\subsection{Bounding}
In order to perform a practical evaluation of programs involving recursive data structures, it is generally necessary to impose some sort of bounding to eliminate the number of lazy initializations to a manageable level. While these bounds may be asserted as invariants within the program in-situ, it is more practical to apply bounding via external means. There are several bounding techniques that may be applied to both GSE and GSESH, including execution path length bounding, k-bounding and n-bounding. k-bounding and n-bounding both operate by imposing limits on the size of the symbolic input heap, but differ in how such bounds are defined. n-bounding limits the number of objects created by lazy initialization, while k-bounding limits the length of initialized reference chains. k-bounding has the property of exhaustively exploring the full breadth of a given data structure, while n-bounding tends to explore deeper heap shapes while being somewhat less thorough in breadth. In addition, k-bounding and n-bounding may be combined to adjust the desired level of approximation between depth and breadth.

While k-bounding and n-bounding are applicable to both GSE and GSESH, they are not both equally suited for comparing the two methods. The main reason for this lies in how how each bounding technique impacts the functional equivalence of the respective methods.  k-bounding preserves the functional equivalence of GSE and GSESH, a fact which can be established by induction over the length of reference chains from a given initial state. In contrast, GSE and GSESH are not generally functionally equivalent under n-bounding. The reason is that since a GSESH state must represent all possible GSE states on an execution path, it generally includes more locations than any single GSE state it represents. This difference in state counts makes it difficult to externally assert n-bounding in a manner that preserves functional equivalence between GSE and GSESH. Functional equivalence may be preserved in the case where n-bounds are asserted via in-situ invariants, but the significant increase in constraint complexity makes this approach impractical. For this reason, n-bounding is was not selected for comparing GSE to GSESH in the tests performed in this paper.

In contrast to the heap-based bounds provided by k-bounding and n-bounding, execution path-length bounding works by limiting the number along any given execution path. As with other bounding techniques, the usefulness of path-length bounding hinges on whether it preserves functional equivalence. While path-length bounding preserves functional equivalence for GSE and GSESH as formulated in this paper, it does not preserve it for the algorithms as they are implemented in JPF. The reason for the discrepancy lies in how states are defined in JPF. In JPF, the notion of states is closely tied to the concept of nondeterminism. New states are created in JPF only when a nondeterministic choice is made, meaning the number of states on an execution path is tied to the number of nondeterministic choices. A discrepancy in the length of the respective execution paths may be caused by any difference in nondeterminism between the two methods. This is significant, since the nondeterminism in GSESH and GSE differs in both field reads and in reference compares. Field reads in GSESH are deterministic, whereas in GSE they may initiate a nondeterministic lazy initialization. Conversely, in GSE address compares are deterministic, but in GSESH address compares require a nondeterministic assertion over the truth of the branch condition. For this reason, path-length bounding was not used for this experimental evaluation. For future tests, we are considering inserting artificial points of nondeterminism in order to establish functional equivalence of path-length bounds between GSE and GSESH.

\subsection{Evaluation} 
Experimental set-up

@@

In order to evaluate the performance of GSESH, the implementation was tested against three examples: LinkedList, BinarySearchTree, TreeMap.
The LinkedList example tests performance on linked list data structures. The test program begins by assuming an object invariant before initiating a sequence of contains() method calls. This is a challenging example for GSE because each call to the contains() method contains a new lazy initialization, which in turn creates an exponential path expansion throughout program execution. While the constraints for each path are easily solved, the sheer number of paths quickly becomes overwhelming.

@@

Like the BinarySearchTree test, the TreeMap test is simply an assertion of an object invariant, in this case for a red/black tree. This example serves as a stress test for GSESH, since most of the nondeterminism comes from linear constraints over the shape of the tree, rather than aliasing constraints created during lazy initialization. In this case, the path advantage for GSESH is minimized.

@@

Analysis

>>>>

\section{Empirical Evaluation}
\symtxt{} creates fewer execution paths than \gsetxt{}, but it does so with increased path constraint complexity. Determining the impact of this tradeoff in real-world applications is an important research question.

Since the number of paths in \gsetxt{} is exponential in the number of lazy initializations, it is expected that programs for which the number of lazy initializations is proportional to program complexity will evaluate more quickly using \symtxt{}. On the other hand, since \symtxt{} places a greater burden on the constraint solver, it is expected that programs where the lazy initialization count is constant or slowly increasing with respect to program complexity will evaluate more quickly using \symtxt{}.

\subsection{JPF Implementation}
The Java Pathfinder (JPF) symbolic execution engine contains an implementation of \gsetxt{}, which we extended with our own implementation of \symtxt{}.

\subsection{Bounding}
In order to perform a practical evaluation of programs involving recursive data structures, it is generally necessary to impose some sort of bounding to eliminate the number of lazy initializations to a manageable level. While these bounds may be asserted as invariants within the program in-situ, it is more practical to apply bounding via external means. There are several bounding techniques that may be applied to both \gsetxt{} and \symtxt{}, including execution path length bounding, k-bounding and n-bounding. k-bounding and n-bounding both operate by imposing limits on the size of the symbolic input heap, but differ in how such bounds are defined. n-bounding limits the number of objects created by lazy initialization, while k-bounding limits the length of initialized reference chains. k-bounding has the property of exhaustively exploring the full breadth of a given data structure, while n-bounding tends to explore deeper heap shapes while being somewhat less thorough in breadth. In addition, k-bounding and n-bounding may be combined to adjust the desired level of approximation between depth and breadth.

While k-bounding and n-bounding are applicable to both \gsetxt{} and \symtxt{}, they are not both equally suited for comparing the two methods. The main reason for this lies in how how each bounding technique impacts the functional equivalence of the respective methods.  k-bounding preserves the functional equivalence of \gsetxt{} and \symtxt{}, a fact which can be established by induction over the length of reference chains from a given initial state. In contrast, \gsetxt{} and \symtxt{} are not generally functionally equivalent under n-bounding. The reason is that since a \symtxt{} state must represent all possible \gsetxt{} states on an execution path, it generally includes more locations than any single \gsetxt{} state it represents. This difference in state counts makes it difficult to externally assert n-bounding in a manner that preserves functional equivalence between \gsetxt{} and \symtxt{}. Functional equivalence may be preserved in the case where n-bounds are asserted via in-situ invariants, but the significant increase in constraint complexity makes this approach impractical. For this reason, n-bounding is was not selected for comparing \gsetxt{} to \symtxt{} in the tests performed in this paper.

In contrast to the heap-based bounds provided by k-bounding and n-bounding, execution path-length bounding works by limiting the number along any given execution path. As with other bounding techniques, the usefulness of path-length bounding hinges on whether it preserves functional equivalence. While path-length bounding preserves functional equivalence for \gsetxt{} and \symtxt{} as formulated in this paper, it does not preserve it for the algorithms as they are implemented in JPF. The reason for the discrepancy lies in how states are defined in JPF. In JPF, the notion of states is closely tied to the concept of nondeterminism. New states are created in JPF only when a nondeterministic choice is made, meaning the number of states on an execution path is tied to the number of nondeterministic choices. A discrepancy in the length of the respective execution paths may be caused by any difference in nondeterminism between the two methods. This is significant, since the nondeterminism in \symtxt{} and \gsetxt{} differs in both field reads and in reference compares. Field reads in \symtxt{} are deterministic, whereas in \gsetxt{} they may initiate a nondeterministic lazy initialization. Conversely, in \gsetxt{} address compares are deterministic, but in \symtxt{} address compares require a nondeterministic assertion over the truth of the branch condition. For this reason, path-length bounding was not used for this experimental evaluation. For future tests, we are considering inserting artificial points of nondeterminism in order to establish functional equivalence of path-length bounds between \gsetxt{} and \symtxt{}.

\subsection{Evaluation} 
Experimental set-up: TO BE ADDED

@@

In order to evaluate the performance of \symtxt{}, the implementation was tested against three examples: LinkedList, BinarySearchTree, TreeMap.
The LinkedList example tests performance on linked list data structures. The test program begins by assuming an object invariant before initiating a sequence of contains() method calls. This is a challenging example for \gsetxt{} because each call to the contains() method contains a new lazy initialization, which in turn creates an exponential path expansion throughout program execution. While the constraints for each path are easily solved, the sheer number of paths quickly becomes overwhelming.

@@

Like the BinarySearchTree test, the TreeMap test is simply an assertion of an object invariant, in this case for a red/black tree. This example serves as a stress test for \symtxt{}, since most of the nondeterminism comes from linear constraints over the shape of the tree, rather than aliasing constraints created during lazy initialization. In this case, the path advantage for \symtxt{} is minimized.

@@

The results of the experiments are shown in table \ref{tab:results}.

\begin{table} [h]
  \centering
  \begin{tabular}{| c | c | c | c | c | c | c | c |}
  \hline
   \multirow{2}{*}{Method }&\multirow{2}{*}{ $k$ }
   &\multicolumn{2}{|c|}{Time} &\multicolumn{2}{|c|}{ Paths }\\
								&	&\gsetxt{}	&SH	&\gsetxt{} & SH\\
   \hline
    \multirow{3}{*}{LinkedList }&3	& 1 & 1  &1656 & 25		 \\
   		 				& 4	& 3 & 2	&17485  & 39 \\
   						& 5	& 20 & 3	&232743 & 56\\
						& 6	& 338 & 8		&3731094 & 76\\
    \hline
    \multirow{3}{*}{BinarySearchTree }&1	& 0 & 1	& 4	 & 4\\
   		 				& 2	& 1 & 1 	& 26 & 17\\
   						& 3	& 12 & 7	& 305 & 118\\
    \hline
      \multirow{3}{*}{TreeMap}&1	& 0 & 1 	&9 & 9 \\
   		 				&2	& 1 & 3		& 100 & 46 \\
   						&3	&21 & 206	& 3026 & 547 \\

    \hline
  \end{tabular}
  \caption{test results}
  \label{tab:results}
\end{table}

\subsection{Analysis}

LinkedList
As expected, the repeated use of lazy initialization throughout the program causes an exponential increase in the number of paths explored by \gsetxt{}. By combining many of the \gsetxt{} paths, SH avoids the exponential path explosion, demonstrating dramatically reduced, though still exponential, execution time growth. Remarkably, path growth in the k-bound appears to be sub-linear.

BST
The BinarySearchTree example is somewhat of a mixed bag.

TreeMap
